---
title: "Data-Project_phase2"
author: "Sander Beerepoot and Tommaso Reali"
format: html
---

Project repository: [GitHub repo](https://github.com/tommasoreali/Data-Management-Final-Project.git)

```{r, message=FALSE}
library(ggplot2)
library(tidyverse)
library(here)
library(vroom)
library(stringr)
library(dplyr)
library(purrr)
library(tibble)
library(tidyr)
library(readr)

```


# Changes made to Health data

## Full data Cleaning with LOOP

```{r}
# Read step:
read_health_file <- function(path) {
  line1 <- vroom::vroom_lines(path, n_max = 1)
  skip_n <- if (stringr::str_detect(line1, stringr::regex("FIPS;State;County", ignore_case = TRUE))) 0 else 1

  vroom::vroom(
    file = path,
    delim = ";",
    skip = skip_n,
    col_names = TRUE,
    col_types = vroom::cols(.default = "c"),  
    show_col_types = FALSE,
    progress = FALSE
  )
}

```


```{r}
# Clean step:

clean_health <- function(df, year) {

  df <- df |>
    dplyr::rename(
      `% Smokers` = dplyr::any_of("% Adults Reporting Currently Smoking"),
      `% Children in Single-Parent Households` = dplyr::any_of("% Single-Parent Households")
    )

  cols_keep <- c(
    "FIPS","State","County",
    "% Severe Housing Problems",
    "Income Ratio",
    "% Unemployed",
    "Mental Health Provider Rate",
    "Teen Birth Rate",
    "% Excessive Drinking",
    "% Smokers",
    "% Children in Single-Parent Households"
  )

  eu <- readr::locale(decimal_mark = ",", grouping_mark = ".")

  df |>
    dplyr::select(dplyr::any_of(cols_keep)) |>
    dplyr::mutate(
      County = dplyr::if_else(is.na(County) | County == "", "Average", County),
      year = year
    ) |>
    dplyr::mutate(dplyr::across(
      -c(FIPS, State, County, year),
      ~ readr::parse_number(.x, locale = eu)
    ))
}

```


```{r, message=FALSE}
#Loop step:

health_files <- list.files(
  path = here::here("Data"),
  pattern = "2020health\\.csv$|2021health\\.csv$|2022health\\.csv$|2023health\\.csv$",
  full.names = TRUE
)

health_total <- tibble::tibble(path = health_files) |>
  dplyr::mutate(
    year = stringr::str_extract(basename(path), "20\\d{2}") |> as.integer(),
    raw = purrr::map(path, read_health_file),
    clean = purrr::map2(raw, year, clean_health)
  ) |>
  dplyr::select(clean) |>
  tidyr::unnest(clean)

```

```{r}
health_total <- health_total |>
  mutate(County = str_c(County, " County")) |>
  rename(Smokers = `% Smokers`, Unemployed = `% Unemployed`,Housing_Problems = `% Severe Housing Problems`, Income_Ratio = `Income Ratio`, Mental_help = `Mental Health Provider Rate`, Teen_births = `Teen Birth Rate`, Child_alone = `% Children in Single-Parent Households`, Excessive_drinking = `% Excessive Drinking`)
  
```

# Changes made for Education data

## Full data Cleaning with LOOP

```{r}
# Read step
read_edu_file <- function(path) {
  first <- vroom_lines(path, n_max = 1)
  delim <- if (str_detect(first, ";")) ";" else ","

  header_tokens <- c("Geographic Area Name", "Geography", "Estimate")
  skip_n <- if (any(str_detect(first, fixed(header_tokens)))) 0 else 1

  vroom(
    file = path,
    delim = delim,
    skip = skip_n,
    col_names = TRUE,
    col_types = vroom::cols(.default = "c"),
    show_col_types = FALSE,
    progress = FALSE
  )
}

```

```{r}
#Clean step
clean_edu <- function(df, year) {
  out <- df |>
    select(
      `Geographic Area Name`,
      `Estimate!!Percent!!AGE BY EDUCATIONAL ATTAINMENT!!Population 25 years and over!!Bachelor's degree or higher`
    ) |>
    rename(
      Bacmore = `Estimate!!Percent!!AGE BY EDUCATIONAL ATTAINMENT!!Population 25 years and over!!Bachelor's degree or higher`
    ) |>
    distinct(`Geographic Area Name`, .keep_all = TRUE) |>
    separate(
      `Geographic Area Name`,
      into = c("County", "State"),
      sep = ",\\s*",
      extra = "merge",
      fill = "right"
    ) |>
    mutate(
      year = year,
      Bacmore = readr::parse_number(
        Bacmore,
        locale = readr::locale(decimal_mark = ".", grouping_mark = ",")
      )
    ) |>
    select(County, State, year, Bacmore)

  out
}

```

```{r}
#Loop Step

edu_files <- list.files(
  path = here("Data"),
  pattern = "Education2020\\.csv$|education2021\\.csv$|education2022\\.csv$|education2023\\.csv$",
  full.names = TRUE
)

edu_total <- tibble(path = edu_files) |>
  mutate(
    year = str_extract(basename(path), "20\\d{2}") |> as.integer(),
    raw  = map(path, read_edu_file),
    clean = map2(raw, year, clean_edu)
  ) |>
  select(clean) |>
  unnest(clean) |>
  select(County, State, year, Bacmore)
```

```{r}
edu_total |>
  arrange(State, County, year )
```

## Merging datasets

```{r}
data_total <- edu_total |>
  left_join(health_total, by = c("County", "State", "year"))
```

## Description of data obtention and cleaning

The raw data for health was at first downloaded from the CHR&R website for all the years we are interested in for this study (2020-2023) as 4 different csv files (one for each year). 

Link: [Health Data](https://www.countyhealthrankings.org/health-data/methodology-and-sources/data-documentation/national-data-documentation-2010-2023)

Similarly, regarding raw data for education it was downloaded from the census bureau by selecting all of the counties as observations and choosing the education attainment as the interest variable. Then 4 different csv files were downloaded corresponding to years 2020-2023.

Link: [Education Data](https://data.census.gov/table?q=S1501:+Educational+Attainment&g=010XX00US\$0500000_040XX00US01\$0500000)

The procedure to create and clean the final health and education dataset is exactly the same for both of them:

Every yearly health/education file is imported using a specific function called: custom reader. This reader initially checks if the dataset contains an extra metadata row before the actual header, because some datasets may have a useless first line and column names starts on second row. If nothing is to be changed this step is skipped. All files are read using a semi column delimiter and all columns are intentionally imported as character strings to prevent automatic misinterpretation by the reader. This ensures the data remains exactly the same as it is initially in the source files.

Furthermore, a second function is used, namely the cleaning function. It first standardized column names which can vary across yearly datasets. Additionally, it only keeps the variables of interest for the analysis and drops the rest, and adds a year identifier for each dataset. Lastly, all non identifier variables are converted from character strings to numeric values by preserving the numerical precision. 

Finally, a loop is used to process all the yearly files and apply to all of them the same reading and cleaning function. Once the data has been imported and cleaned across all datasets, we merge them and form a single long format dataset (one for Education and one for Health). With this method we avoid redundant code and consistency across years is garanteed. 

After having done this procedure for the Health and the Education datasets, they can be merged together by using County, State and Year as keys to create the final dataset used in this study.

## Description of variables

variables of Interest

- Adult smoking (in percentage of total of population of adults): This variable describes the reported rate of current daily smokers in the population of the county.

- Excessive drinking: This variable describes the self-reported rate of excessive drinkers of the population in the county.

- Teen births: This variable describes the amount of teen births per 1000 females aged 15-19.

Main variable

- Education level (percent of 25 > year olds that have a Bachelors degree minimum): This variable describes the percentage of 25 year olds or older that have a diploma for a Bachelors degree or higher. 

Control variables to complete model: 

- Unemployment: This variable describes the unemployment rate in the county. Unemployed people are 16 years old or older and are actively looking for work. 

- Income Inequality: This variable describes the income inequality levels, computed as the ratio of income at the 80th percentile over the income at the 20th percentile. 

- Severe Housing Problems: this variable describes the percentage of households that have at least one of the 4 reported housing problems: high housing costs, overcrowding, lack of kitchen or plumbing facilities.

- Children in single parent households: This variable describes the percentage of children that grow up in single parent households currently.

- Mental health providers: This variable describes the rate of mental health providers per 100000 inhabitants. 


## Research question description (500 words):

The topic we are going to analyse focuses on studying different health habits based on the educational attainment in an observed area, in this case, a county. The variables that will be looked into are adult smoking, excessive drinking, and teen births per 1000 15-19 year old women. We have chosen these variables because these are in someway based on choices that individuals make, rather than an unlucky occurrence in someones life. This means that, for example, getting sick can be the result of bad choices, however someone can not decide to become sick from one day to the other. On the contrary, choosing to smoke, choosing to excessively drink or delivering a baby during teen years are more closely related to choices and thus the environment. This is why we believe it is important to look into the determinants of these habits. One of the possible determinants is educational attainment because it not only influences individuals in their own lives, it also influences individuals in a community. Educational Attainment is computed as the proportion of people in a county that have obtained a Bachelors degree at least. The goal will be to analyse the aggregate decisions of individuals on health habits based on the aggregated educational attainment in a county and compare between more and less educated counties to see whether there are disparities based on educational attainment. 
Maralani V. (2014) advances that education on its own does not lead to better health, rather, the bundle of educational inequality and environment plays a much larger role in predicting whether someone will smoke. 
This research will thus try and adapt these findings further on smoking and adapting it to other health habits. 
The estimation will be completed with different control variables, which we believe can isolate the impact of educational attainment. these are: - Children in single parent households
- mental health providers
- Severe Housing Problems
- Income inequality
- Unemployment
The vast amount of counties in this dataset, spread through the 51 states of the USA, allow for a large sample size, and thus the development of different conclusions, whether some choices are regionally based, due to a certain education level, or completely random.  


## Description of the sources: 

The education data set comes from the United States Census Bureau. It serves as the main national statistics agency in the US. It is a national data organization that serves the mission to be the US's leading provider of data about the US citizens and the country's economy. The bureau collects a varying range of data on businesses, education, and housing. The datasets which we have chosen from the US census bureau are the datasets relating to the educational attainment, 5-year ACS estimates, for the years 2020-2023. These datasets provides data on education across the counties of the US, dividing in different groups based on age, ethnicity, and poverty rate if needed. 


The Health dataset used in this study is constructed of data coming entirely from the CHR&R (County Health Rankings and Roadmaps) program. This initiative was created by the University of Wisconsin Population Health institute with support from the Robert Wood Johnson Foundation. It consists in providing annual data on health factors within and across almost the entirety of counties from the United States. The CHR&R provides data, evidence, guidance and examples to build awareness about the multiple factors influencing health conditions and disparities across the country. Furthermore, CHR&R puts an emphasis on the range of health-influencing measures that offer crucial context about community specific health conditions. In fact, Along with the data the program provides tools and evidence-based strategies, which can be used by the communities to move to action and solve the important health and societal issues concerning their people. As an example, The "2025 Annual data release" includes over 80 health-related measures, and focuses on the impact of certain societal factors present in the United States on community conditions.

# Tables per dataset

## Table of education data set

```{r}
data_name <- "Education Data"

row_count <- nrow(EDU2Long)
column_count <- ncol(EDU2Long)

summary_table <- data.frame(
  `Data File Name` = data_name,
  `Number of Rows` = row_count,
  `Number of Columns` = column_count,
  check.names = FALSE
)
print(summary_table)
```

## Table of health data set

```{r}
data_name <- "Health Data"

row_count <- nrow(Healthtotal)
column_count <- ncol(Healthtotal)

summary_table <- data.frame(
  `Data File Name` = data_name,
  `Number of Rows` = row_count,
  `Number of Columns` = column_count,
  check.names = FALSE
) |>
  knitr::kable()
```

# Tables for Main Variables

## Table for Education variable

Since there are 3144 counties, the data will be aggregated. 

```{r, message=FALSE}
EDU2Long|>
  ggplot(aes(y = Bacmore, x = Year)) +
  geom_violin()
```
While not the most accurate, these violin plots show the density of educational attainment aggregated in the US. The graphs show that every year, the average bachelors obtention increases slightly, going from about 17% to 20% between 2020 and 2023. 

## Table for health main target variables (% Smokers, % Excessive drinkers, Teen birth rates)

```{r, message=FALSE}
Healthtotal|>
  ggplot(aes(y = TeenBirth, x = Year)) +
  geom_violin()
```

These violin plots show the aggregated US average of Teen birth rates as the amount of teen births per 1000 females aged 15-19. It is observed that the density of the violin plots gets more and more concentrated towards the bottom as the years pass by. As a matter of fact, the average of aggregated teen birth lowers from about 25% in 2020 to 19% in 2023.

```{r, message=FALSE}
Healthtotal|>
  ggplot(aes(y = Smokers, x = Year)) +
  geom_violin()
```

These violin plots show the aggregated US average of adult smokers as a percentage of the population. It can be seen that each year the average of percentage of smokers gradually increases and the density of the violin plots gets slimmer. In fact, in 2020 the average percentage of aggregated smokers in the US is about just above 15%. It grows up until nearly 20% in 2021, and decreases to about 18% in 2023.

```{r, message=FALSE}

Healthtotal|>
  ggplot(aes(y = ExcessiveDrink, x = Year)) +
  geom_violin()
```

These violin plots showing the aggregated US average of Excessive drinking as a percentage of the population highlight a thicker upper tail as the years pass by, meaning there is a higher count of people excessively drinking. However the average of the percentage of excessive drinkers doesn't seem to change from from 2020 to 2023.



