---
title: "PHASE-3"
author: 
  Tommaso Reali 
  Sander Beerepot 
format:
  html: default
  pdf:  default
editor: visual
execute: 
  warning: false
  echo: false
filters: 
- wordcount
---

Project repository: [GitHub repo](https://github.com/tommasoreali/Data-Management-Final-Project.git)

```{r}
library(ggplot2)
library(tidyverse)
library(here)
library(vroom)
library(stringr)
library(dplyr)
library(purrr)
library(tibble)
library(tidyr)
library(readr)

```


# Research question

The topic we are going to analyse focuses on studying different health habits based on the educational attainment in an observed area, in this case, a county. The variables that will be looked into are adult smoking, excessive drinking, and teen births per 1000 15-19 year old women. We have chosen these variables because these are in someway based on choices that individuals make, rather than an unlucky occurrence in someones life. This means that, for example, getting sick can be the result of bad choices, however someone can not decide to become sick from one day to the other. On the contrary, choosing to smoke, choosing to excessively drink or delivering a baby during teen years are more closely related to choices and thus the environment. This is why we believe it is important to look into the determinants of these habits. One of the possible determinants is educational attainment because it not only influences individuals in their own lives, it also influences individuals in a community. Educational Attainment is computed as the proportion of people in a county that have obtained a Bachelors degree at least. The goal will be to analyse the aggregate decisions of individuals on health habits based on the aggregated educational attainment in a county and compare between more and less educated counties to see whether there are disparities based on educational attainment. Maralani V. (2014) advances that education on its own does not lead to better health, rather, the bundle of educational inequality and environment plays a much larger role in predicting whether someone will smoke. This research will thus try and adapt these findings further on smoking and adapting it to other health habits. The estimation will be completed with different control variables, which we believe can isolate the impact of educational attainment. these are: - Children in single parent households - mental health providers - Severe Housing Problems - Income inequality - Unemployment The vast amount of counties in this dataset, spread through the 51 states of the USA, allow for a large sample size, and thus the development of different conclusions, whether some choices are regionally based, due to a certain education level, or completely random.

# Datasets description

```{r}
# Creation of the final Health Dataset:
# Read step:
read_health_file <- function(path) {
  line1 <- vroom::vroom_lines(path, n_max = 1)
  skip_n <- if (stringr::str_detect(line1, stringr::regex("FIPS;State;County", ignore_case = TRUE))) 0 else 1

  vroom::vroom(
    file = path,
    delim = ";",
    skip = skip_n,
    col_names = TRUE,
    col_types = vroom::cols(.default = "c"),  
    show_col_types = FALSE,
    progress = FALSE
  )
}

```


```{r}
# Creation of the final Health Dataset:
# Clean step:

clean_health <- function(df, year) {

  df <- df |>
    dplyr::rename(
      `% Smokers` = dplyr::any_of("% Adults Reporting Currently Smoking"),
      `% Children in Single-Parent Households` = dplyr::any_of("% Single-Parent Households")
    )

  cols_keep <- c(
    "FIPS","State","County",
    "% Severe Housing Problems",
    "Income Ratio",
    "% Unemployed",
    "Mental Health Provider Rate",
    "Teen Birth Rate",
    "% Excessive Drinking",
    "% Smokers",
    "% Children in Single-Parent Households"
  )

  eu <- readr::locale(decimal_mark = ",", grouping_mark = ".")

  df |>
    dplyr::select(dplyr::any_of(cols_keep)) |>
    dplyr::mutate(
      County = dplyr::if_else(is.na(County) | County == "", "Average", County),
      year = year
    ) |>
    dplyr::mutate(dplyr::across(
      -c(FIPS, State, County, year),
      ~ readr::parse_number(.x, locale = eu)
    ))
}

```


```{r}
# Creation of the final Health Dataset:
#Loop step:

health_files <- list.files(
  path = here::here("Data"),
  pattern = "2020health\\.csv$|2021health\\.csv$|2022health\\.csv$|2023health\\.csv$",
  full.names = TRUE
)

health_total <- tibble::tibble(path = health_files) |>
  dplyr::mutate(
    year = stringr::str_extract(basename(path), "20\\d{2}") |> as.integer(),
    raw = purrr::map(path, read_health_file),
    clean = purrr::map2(raw, year, clean_health)
  ) |>
  dplyr::select(clean) |>
  tidyr::unnest(clean)

```

```{r}
health_total <- health_total |>
  mutate(County = str_c(County, " County")) |>
  rename(Smokers = `% Smokers`, Unemployed = `% Unemployed`,Housing_Problems = `% Severe Housing Problems`, Income_Ratio = `Income Ratio`, Mental_help = `Mental Health Provider Rate`, Teen_births = `Teen Birth Rate`, Child_alone = `% Children in Single-Parent Households`, Excessive_drinking = `% Excessive Drinking`)
  
```

The raw data for health was at first downloaded from the CHR&R website for all the years we are interested in for this study (2020-2023) as 4 different csv files (one for each year). 

Link: [Health Data](https://www.countyhealthrankings.org/health-data/methodology-and-sources/data-documentation/national-data-documentation-2010-2023)

Similarly, regarding raw data for education it was downloaded from the census bureau by selecting all of the counties as observations and choosing the education attainment as the interest variable. Then 4 different csv files were downloaded corresponding to years 2020-2023.

Link: [Education Data](https://data.census.gov/table?q=S1501:+Educational+Attainment&g=010XX00US\$0500000_040XX00US01\$0500000)

The procedure to create and clean the final health and education dataset is exactly the same for both of them:

Every yearly health/education file is imported using a specific function called: custom reader. This reader initially checks if the dataset contains an extra metadata row before the actual header, because some datasets may have a useless first line and column names starts on second row. If nothing is to be changed this step is skipped. All files are read using a semi column delimiter and all columns are intentionally imported as character strings to prevent automatic misinterpretation by the reader. This ensures the data remains exactly the same as it is initially in the source files.

Furthermore, a second function is used, namely the cleaning function. It first standardized column names which can vary across yearly datasets. Additionally, it only keeps the variables of interest for the analysis and drops the rest, and adds a year identifier for each dataset. Lastly, all non identifier variables are converted from character strings to numeric values by preserving the numerical precision. 

Finally, a loop is used to process all the yearly files and apply to all of them the same reading and cleaning function. Once the data has been imported and cleaned across all datasets, we merge them and form a single long format dataset (one for Education and one for Health). With this method we avoid redundant code and consistency across years is garanteed. 

After having done this procedure for the Health and the Education datasets, they can be merged together by using County, State and Year as keys to create the final dataset used in this study. 


```{r}
# Creation of the final Education Dataset:
# Read step
read_edu_file <- function(path) {
  first <- vroom_lines(path, n_max = 1)
  delim <- if (str_detect(first, ";")) ";" else ","

  header_tokens <- c("Geographic Area Name", "Geography", "Estimate")
  skip_n <- if (any(str_detect(first, fixed(header_tokens)))) 0 else 1

  vroom(
    file = path,
    delim = delim,
    skip = skip_n,
    col_names = TRUE,
    col_types = vroom::cols(.default = "c"),
    show_col_types = FALSE,
    progress = FALSE
  )
}

```

```{r}
# Creation of the final Education Dataset:
# Clean step
clean_edu <- function(df, year) {
  out <- df |>
    select(
      `Geographic Area Name`,
      `Estimate!!Percent!!AGE BY EDUCATIONAL ATTAINMENT!!Population 25 years and over!!Bachelor's degree or higher`
    ) |>
    rename(
      Bacmore = `Estimate!!Percent!!AGE BY EDUCATIONAL ATTAINMENT!!Population 25 years and over!!Bachelor's degree or higher`
    ) |>
    distinct(`Geographic Area Name`, .keep_all = TRUE) |>
    separate(
      `Geographic Area Name`,
      into = c("County", "State"),
      sep = ",\\s*",
      extra = "merge",
      fill = "right"
    ) |>
    mutate(
      year = year,
      Bacmore = readr::parse_number(
        Bacmore,
        locale = readr::locale(decimal_mark = ".", grouping_mark = ",")
      )
    ) |>
    select(County, State, year, Bacmore)

  out
}

```

```{r}
# Creation of the final Education Dataset:
# Loop Step

edu_files <- list.files(
  path = here("Data"),
  pattern = "Education2020\\.csv$|education2021\\.csv$|education2022\\.csv$|education2023\\.csv$",
  full.names = TRUE
)

edu_total <- tibble(path = edu_files) |>
  mutate(
    year = str_extract(basename(path), "20\\d{2}") |> as.integer(),
    raw  = map(path, read_edu_file),
    clean = map2(raw, year, clean_edu)
  ) |>
  select(clean) |>
  unnest(clean) |>
  select(County, State, year, Bacmore)
```

```{r}
data_total <- edu_total |>
  left_join(health_total, by = c("County", "State", "year"))
```



```{r}
here::i_am("Data-Management-Final-Project.Rproj")
```

```{r, message=FALSE, echo=FALSE}
health2020 <- vroom(here("Data", "2020health.csv"))
health2021 <- vroom(here("Data", "2021health.csv"))
health2022 <- vroom(here("Data", "2022health.csv"))
health2023 <- vroom(here("Data", "2023health.csv"))
education2020 <- vroom(here("Data", "Education2020.csv"))
education2021 <- vroom(here("Data", "Education2021.csv"))
education2022 <- vroom(here("Data", "Education2022.csv"))
education2023 <- vroom(here("Data", "Education2023.csv"))
```


```{r}

datasets <- list(
  edu    = edu_total,
  health = health_total,
  total   = data_total
)

basic_info <- imap_dfr(datasets, ~{
  df <- .x
  tibble(
    file = .y,
    row_number = nrow(df),
    col_number = ncol(df),
    missing_number = sum(is.na(df)),
    rows_with_missing_values = sum(rowSums(is.na(df)) > 0)
  )
})

knitr::kable(basic_info, digits = 0, caption = "Basic information about the data files"
)

```

# Description of the sources

The education data set comes from the United States Census Bureau. It serves as the main national statistics agency in the US. It is a national data organization that serves the mission to be the US's leading provider of data about the US citizens and the country's economy. The bureau collects a varying range of data on businesses, education, and housing. The datasets which we have chosen from the US census bureau are the datasets relating to the educational attainment, 5-year ACS estimates, for the years 2020-2023. These datasets provides data on education across the counties of the US, dividing in different groups based on age, ethnicity, and poverty rate if needed.

The Health dataset used in this study is constructed of data coming entirely from the CHR&R (County Health Rankings and Roadmaps) program. This initiative was created by the University of Wisconsin Population Health institute with support from the Robert Wood Johnson Foundation. It consists in providing annual data on health factors within and across almost the entirety of counties from the United States. The CHR&R provides data, evidence, guidance and examples to build awareness about the multiple factors influencing health conditions and disparities across the country. Furthermore, CHR&R puts an emphasis on the range of health-influencing measures that offer crucial context about community specific health conditions. In fact, Along with the data the program provides tools and evidence-based strategies, which can be used by the communities to move to action and solve the important health and societal issues concerning their people. As an example, The "2025 Annual data release" includes over 80 health-related measures, and focuses on the impact of certain societal factors present in the United States on community conditions.

# Description of variables

variables of Interest:

-   Adult smoking (in percentage of total of population of adults): This variable describes the reported rate of current daily smokers in the population of the county, and is labelled as "Smokers".

-   Excessive drinking: This variable describes the self-reported rate of excessive drinkers of the population in the county, and is labelled as "Excessive_drinking".

-   Teen births: This variable describes the amount of teen births per 1000 females aged 15-19, and is labelled as "Teen_births".

Main variable:

-   Education level (percent of 25 \> year olds that have a Bachelors degree minimum): This variable describes the percentage of 25 year olds or older that have a diploma for a Bachelors degree or higher. This variable is labelled as "Bacmore" in this study

Control variables to complete model:

-   Unemployment: This variable describes the unemployment rate in the county. Unemployed people are 16 years old or older and are actively looking for work. It is labelled as "Ünemployed"

-   Income Inequality: This variable describes the income inequality levels, computed as the ratio of income at the 80th percentile over the income at the 20th percentile. It is labelled as "Income_Ratio"

-   Severe Housing Problems: this variable describes the percentage of households that have at least one of the 4 reported housing problems: high housing costs, overcrowding, lack of kitchen or plumbing facilities. This variable is labelled as "Housing_Problems"

-   Children in single parent households: This variable describes the percentage of children that grow up in single parent households currently, and is labelled as "Child_alone".

-   Mental health providers: This variable describes the rate of mental health providers per 100000 inhabitants, and is labelled as "Mental_help".


# Violin plots for Main Variables

Since there are 3144 counties, the data will be aggregated for the implementation of these violin plots.

```{r}
edu_total|>
  ggplot(aes(y = Bacmore, x = year, group= year)) +
  geom_violin() +
labs(
    title = "Distribution of educational attainment by year",
    x = "Year",
    y = "Share with Bachelor's degree or higher (%)"
  )
```

While not the most accurate, these violin plots show the density of educational attainment aggregated in the US. The graphs show that every year, the average bachelors obtention increases slightly, going from about 17% to 20% between 2020 and 2023.

```{r}
health_total|>
  ggplot(aes(y = Teen_births, x = year, group = year)) +
  geom_violin() +
labs(
    title = "Distribution of Teen births by year",
    x = "Year",
    y = "Teen Births per 1000 females"
  )
```

These violin plots show the aggregated US average of Teen birth rates as the amount of teen births per 1000 females aged 15-19. It is observed that the density of the violin plots gets more and more concentrated towards the bottom as the years pass by. As a matter of fact, the average of aggregated teen birth lowers from about 25% in 2020 to 19% in 2023.

```{r}
health_total|>
  ggplot(aes(y = Smokers, x = year, group = year)) +
  geom_violin() +
labs(
    title = "Distribution of Smokers by year",
    x = "Year",
    y = "share of smokers (%)"
  )
```

These violin plots show the aggregated US average of adult smokers as a percentage of the population. It can be seen that each year the average of percentage of smokers gradually increases and the density of the violin plots gets slimmer. In fact, in 2020 the average percentage of aggregated smokers in the US is about just above 15%. It grows up until nearly 20% in 2021, and decreases to about 18% in 2023.

```{r}

health_total|>
  ggplot(aes(y = Excessive_drinking, x = year, group = year)) +
  geom_violin()+
labs(
    title = "Distribution of Excessive drinkers by year",
    x = "Year",
    y = "share of Excessive drinkers (%)"
  )
```

These violin plots showing the aggregated US average of Excessive drinking as a percentage of the population highlight a thicker upper tail as the years pass by, meaning there is a higher count of people excessively drinking. However the average of the percentage of excessive drinkers doesn't seem to change from 2020 to 2023.

# Data Analysis

## Descriptive statistics

```{r}

desc_stats <- data_total |>
  summarise(
    across(
      c(
        Teen_births,
        Smokers,
        Excessive_drinking,
        Bacmore,
        Housing_Problems,
        Income_Ratio,
        Mental_help,
        Child_alone,
        Unemployed
      ),
      ~c(
        Mean = mean(.x, na.rm = TRUE),
        SD   = sd(.x, na.rm = TRUE),
        Min  = min(.x, na.rm = TRUE),
        Max  = max(.x, na.rm = TRUE),
        Median = median(.x, na.rm = TRUE)
      )
    )
  )

desc_table <- as.data.frame(t(desc_stats))
colnames(desc_table) <- c("Mean", "SD", "Min", "Max", "Median")

knitr::kable(desc_table, caption = "Descriptive statistics of main variables")

```


The above descriptive statistic table shows us that there is high heterogeneity observed across counties with wide ranges and high dispersion of data, especially for variables like Teen births and Mental help. Additionally, this is also the case for the educational attainment variable "Bacmore" as it holds values ranging from 0 % to almost 80 %. In fact this type of distribution for educational attainment justifies the use of quartiles for education in subsequent analyses. Furthermore it has a mean of ~23.31 % and a lower median of 20.9%, justifying the distribution is right skewed (which is also the case for other variables like Mental_help). However, behavioral health variables such as Smokers and Excessive_drinking have lower dispersion, entailing systematic difference across counties, a more even distribution and extreme outliers less present in the data. Finally, the socioeconomic control variables such as : Income_Ratio, Housing_problems and Unemployment also vary substantially and have plausible ranges. 

## Within-state deviation analysis

```{r}

# Education quartiles
data_grad <- data_total |>
  mutate(
    edu_q = ntile(Bacmore, 4),
    edu_q = factor(edu_q, levels = 1:4, labels = c("Q1", "Q2", "Q3", "Q4"))
  )

# Compute within-state deviations
within_state <- data_grad |>
  group_by(State, year) |>
  mutate(
    Teen_births_dev = Teen_births - mean(Teen_births, na.rm = TRUE),
    Smokers_dev = Smokers - mean(Smokers, na.rm = TRUE),
    Excessive_drinking_dev = Excessive_drinking - mean(Excessive_drinking, na.rm = TRUE)
  ) |>
  ungroup() |>
  select(edu_q, Teen_births_dev, Smokers_dev, Excessive_drinking_dev) |>
  pivot_longer(
    cols = ends_with("_dev"),
    names_to = "Outcome",
    values_to = "Deviation"
  ) |>
  mutate(
    Outcome = recode(
      Outcome,
      Teen_births_dev = "Teen births",
      Smokers_dev = "Smokers (%)",
      Excessive_drinking_dev = "Excessive drinking (%)"
    )
  )

# Plot: deviations by education quartile
ggplot(within_state, aes(x = edu_q, y = Deviation)) +
  geom_violin(fill = "beige", color = "grey30") +
  geom_boxplot(width = 0.12, outlier.shape = NA) +
  facet_wrap(~Outcome, scales = "free_y") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  labs(
    title = "Within-state deviation analysis",
    x = "Education quartile (within state)",
    y = "Deviation from state mean"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    strip.text = element_text(face = "bold")
  )

```

This within state deviation analysis shown in the above violin plots illustrates how the studied health outcomes (at county level) differ from their state mean for each of the 4 quartiles. Each Violin plot represents the distribution of counties deviations from the mean state for a health outcome, conditional on belonging to a specific quartile (the data has been grouped by state and year). We set "Deviation from state mean" = county's value - state mean value. A deviation of 0 means a county's outcome is equal to its state average. When the deviation is positive (negative) it indicates a worse (better) outcome relative to its state average value. This analysis is helpful as we are comparing counties inside the same State, and we therefore control for state-level policies, institutions and factors. When focusing on the health variable "Excessive drinking", as the education quartile increases "Excessive drinking" deviation from mean state has larger positive values. This entails, that within the same state, counties with higher education levels tend to show higher excessive drinking rates compared to their state average. When focusing on the other two health variables: Smokers % and Teen births the opposite pattern occurs, namely counties with higher education level are associated with lower smoking rates and lower teen birth counts compared to their state mean value. 

## Education gradients
```{r}

edu_trend <- data_grad |>
  group_by(year, edu_q) |>
  summarise(
    Teen_births = mean(Teen_births, na.rm = TRUE),
    Smokers = mean(Smokers, na.rm = TRUE),
    Excessive_drinking = mean(Excessive_drinking, na.rm = TRUE),
    .groups = "drop"
  ) |>
  pivot_longer(
    cols = c(Teen_births, Smokers, Excessive_drinking),
    names_to = "Outcome",
    values_to = "Mean"
  )

ggplot(edu_trend, aes(x = year, y = Mean, color = edu_q, group = edu_q)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  facet_wrap(~Outcome, scales = "free_y") +
  labs(
    title = "Trends over time by education quartile",
    x = "Year",
    y = "Average value",
    color = "Education quartile"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "right",
    panel.grid.minor = element_blank(),
    strip.text = element_text(face = "bold")
  )

```

This education gradient analysis follows the trends first observed and discussed in the previous violin plots of the within state deviation analysis. These graphs reveal the evolution of the health habits per education quartile over the 4 years we study. For excessive drinking, we can see that all quartiles on average follow the same pattern, except for the 1st quartile. a sharp increase in 2021, which may be due to covid, then slight decreases. For the first quartile, we can see that it even slightly increases after 2021. The more educated a county is, the higher the rate of excessive drinkers on average. For the smokers part, we can distinguish the same pattern as for excessive drinking, meaning a sharp increase in 2021, followed by a slight decreases in 2022 and 2023. As opposed to excessive drinking, higher smoking rates are ususally associated with lower education levels. Finally, regarding teen births, there is a general trend showing a decrease between 2020 and 2022, with a slight increase in 2023. The higher teen birth rates are associated with lower education levels.

## Scatterplots

```{r}

scatter_df <- data_total |>
  select(Bacmore, Teen_births, Smokers, Excessive_drinking) |>
  pivot_longer(
    cols = c(Teen_births, Smokers, Excessive_drinking),
    names_to = "Outcome",
    values_to = "Value"
  ) |>
  mutate(
    Outcome = recode(
      Outcome,
      Teen_births = "Teen births",
      Smokers = "Smokers (%)",
      Excessive_drinking = "Excessive drinking (%)"
    )
  )

ggplot(scatter_df, aes(x = Bacmore, y = Value)) +
  geom_point(alpha = 0.15) +
  geom_smooth(method = "lm", se = TRUE) +
  facet_wrap(~Outcome, scales = "free_y") +
  labs(
    x = "Share with Bachelor's degree or more (Bacmore)",
    y = "Outcome value"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    strip.text = element_text(face = "bold")
  )
```

These scatterplots add to what was previously stated. In fact, there is a positive correlation between education and drinking. Furthermore, there is also the negative correlation between smokers and education and between teen births and education.

## Correlation matrix

```{r}

# Correlation matrix

corr_df <- data_total |>
  select(
    Teen_births,
    Smokers,
    Excessive_drinking,
    Bacmore,
    Housing_Problems,
    Income_Ratio,
    Unemployed,
    Mental_help,
    Child_alone
  ) |>
  drop_na()  

corr_mat <- cor(corr_df, use = "complete.obs")

corr_long <- as.data.frame(as.table(corr_mat)) |>
  rename(Var1 = Var1, Var2 = Var2, Corr = Freq)

ggplot(corr_long, aes(x = Var1, y = Var2, fill = Corr)) +
  geom_tile() +
  coord_equal() +
  labs(x = NULL, y = NULL, fill = "Corr.") +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid = element_blank()
  )

```

The above correlation matrix includes all the variables used in this analysis. It can be seen that health outcomes such as Teen birth rates and percentage of Smokers are strongly positively correlated with socioeconomic indicators like income inequality, unemployment and the share of children living in single-parent households. On the other hand, the percentage of excessive drinkers is strongly negatively correlated with Income ratio, percentage of Unemployed and the share of children living in single-parent households. The highlighted correlations suggest the shown health outcomes are probably embedded within broad socioeconomic contexts. Furthermore, several of the socioeconomic indicators are closely related to educational attainment. Therefore, not including the socioeconomic indicators as control variables within the regressions would imply a risk of "Omitted Variable Bias". These controls are then chosen for this regression analysis to assess if education level effect on health outcomes persists once these related factors are taken into account.



```{r}

data_total <- 
  left_join(edu_total, health_total, by = c("County", "State", "year"))
  
```

## Regressions

The analysis and interpretation of the before shown graphs and plots states a clear direction between educational attainment and each specific health outcome. To confirm these trends and to mostly understand if we can talk about an actual impact of a variable on another, we will derive 3 main Regressions.
For interesting data analysis results, the regressions are run in different steps, first independent on dependent variable, then adding socioeconomic variables, then add the other variables. This way we can determine the general impact of each variable and how the effect of the independent variable takes from the other variables.

### Regressions with only explanatory variable

```{r}
simple_smokers <- lm(Smokers ~ Bacmore, data = data_total)
summary(simple_smokers)
simple_teens <- lm(Teen_births ~ Bacmore, data = data_total)
summary(simple_teens)
simple_drinking <- lm(Excessive_drinking ~ Bacmore, data = data_total)
summary(simple_drinking)
```


These regressions are not interesting for causal estimation because they do not include and=y control variables and the coefficients are most probably biased. However, they are interesting to determine the general impact of education on our three dependent variables. It can be determined that education increases excessive drinking habits whilst when education increases, smoking and the amount of children born from teenage mothers decrease. All coefficients are statistically significant

### Regressions including all control variables but without fixed effects

```{r}
regression_smoking <- lm(Smokers ~ Bacmore + Housing_Problems + Income_Ratio + Mental_help + Child_alone + Unemployed, data = data_total)
summary(regression_smoking)

regression_births <- lm(Teen_births ~ Bacmore + Housing_Problems + Income_Ratio + Mental_help + Child_alone + Unemployed, data = data_total)
summary(regression_births)

regression_drinking <- lm(Excessive_drinking ~ Bacmore + Housing_Problems + Income_Ratio + Mental_help + Child_alone + Unemployed, data = data_total)
summary(regression_drinking)
```

this document has {{wordcount}} words
