---
title: "Sander Quarto Doc"
format: html
---
```{r}
library(here)
library(ggplot2)
library(dplyr)
library(vroom)
library(tidyr)
library(tidyverse)
```

##Dowload data

```{r}
here::i_am("Data-Management-Final-Project.Rproj")
```

Description of how to get the data sets

To get the data of the CHR&R, download each file per year seperately

Link: "https://www.countyhealthrankings.org/health-data/methodology-and-sources/data-documentation/national-data-documentation-2010-2023"

To get the education data from the census bureau: Select all of the counties as variables, pick the variable you want to pick (in this case pick the education attainment) and download the data for all the years you are interested in

Link: "https://data.census.gov/table?q=S1501:+Educational+Attainment&g=010XX00US\$0500000_040XX00US01\$0500000"

```{r}
health2020 <- vroom(here("Data", "2020health.csv"))
health2021 <- vroom(here("Data", "2021health.csv"))
health2022 <- vroom(here("Data", "2022health.csv"))
health2023 <- vroom(here("Data", "2023health.csv"))
education2020 <- vroom(here("Data", "ACSST5Y2020.S1501-Data.csv"))
education2021 <- vroom(here("Data", "ACSST5Y2021.S1501-Data.csv"))
education2022 <- vroom(here("Data", "ACSST5Y2022.S1501-Data.csv"))
education2023 <- vroom(here("Data", "ACSST5Y2023.S1501-Data.csv"))

```

## Changes made to health data

### Changes for health2020

```{r}
#replace wrong column names with first row:
new_names <- as.character(health2020[1, ])
colnames(health2020) <- new_names
health2020 <- health2020[-1, ]

```

```{r}
#remove duplicates and their first occurence because we don't care about them
names(health2020) <- make.names(names(health2020), unique = TRUE)
to_drop <- grep("\\.[0-9]+$", names(health2020), value = TRUE)
health2020 <- health2020 %>% select(-all_of(to_drop))

```

```{r}
#change character columns in numeric
health2020 <- health2020 %>%
  mutate(across(
    -c(State, County),   # keep these as character
    ~ suppressWarnings(as.numeric(.x))
  ))
```

## Changes made for Education data

###changes for education2020

```{r}
#replace wrong column names with first row:
new_names_education <- as.character(education2020[1, ])
colnames(education2020) <- new_names_education
education2020 <- education2020[-1, ]
```


```{r}
education2020 <- education2020 %>%
  select(-contains("Margin of Error", ignore.case = TRUE))|>
  select(-contains("Total", ignore.case = TRUE))|>
  select(-contains("Estimate!!Male!!", ignore.case = TRUE))|>
  select(-contains("Estimate!!Female!!", ignore.case = TRUE))|>
  select(-contains("High school", ignore.case = TRUE)) |>
  select(-contains("Graduate", ignore.case = TRUE))|>
  select(-contains("18 to 24", ignore.case = TRUE))|>
  select(-contains("34", ignore.case = TRUE))|>
  select(-contains("9th", ignore.case = TRUE))|>
  select(-contains("34", ignore.case = TRUE))|>
  select(-contains("Associate's", ignore.case = TRUE))|>
  select(-contains("34", ignore.case = TRUE))|>
  select(-contains("Population 65 years and over", ignore.case = TRUE))|>
  select(-contains("Population 45 to 64 years", ignore.case = TRUE))|>
  select(-contains("Population 35 to 44 years", ignore.case = TRUE))|>
  select(-contains("RACE", ignore.case = TRUE))|>
  select(-contains("college", ignore.case = TRUE))
```



```{r}
#easier code for directly deleting all the rows in a single code 
education2020 <- education2020 %>%
  select(
    -matches("Margin of Error|Total|Estimate!!Male!!|Estimate!!Female!!",
             ignore.case = TRUE)
  )
```

```{r}
# remove columns that are ONLY "(X)" (ignoring NAs)
education2020 <- education2020 %>%
  select(
    where(~ !all(.x == "(X)" | is.na(.x)))
  )
```


```{r}
#remove also percent of females and percent of males in this column (if you want), this code can be kept or deleted depending on our decision to keep Percentage Female, PErcentage Male.
education2020 <- education2020 %>%
  select(
    -matches("Estimate!!Percent Male!!|Estimate!!Percent Female!!",
             ignore.case = TRUE)
  )
```

```{r}
#convert to numeric columns the ones that need it

education2020 <- education2020 %>%
  mutate(across(
    -c(Geography, `Geographic Area Name`),
    ~ suppressWarnings(as.numeric(.x))
  ))

```

## Changes for education 2021

```{r}
#replace wrong column names with first row:
new_names_education <- as.character(education2021[1, ])
colnames(education2021) <- new_names_education
education2021 <- education2021[-1, ]
```


```{r}
education2021 <- education2021 %>%
  select(-contains("Margin of Error", ignore.case = TRUE))|>
  select(-contains("Total", ignore.case = TRUE))|>
  select(-contains("Estimate!!Male!!", ignore.case = TRUE))|>
  select(-contains("Estimate!!Female!!", ignore.case = TRUE))|>
  select(-contains("High school", ignore.case = TRUE)) |>
  select(-contains("Graduate", ignore.case = TRUE))|>
  select(-contains("18 to 24", ignore.case = TRUE))|>
  select(-contains("34", ignore.case = TRUE))|>
  select(-contains("9th", ignore.case = TRUE))|>
  select(-contains("34", ignore.case = TRUE))|>
  select(-contains("Associate's", ignore.case = TRUE))|>
  select(-contains("34", ignore.case = TRUE))|>
  select(-contains("Population 65 years and over", ignore.case = TRUE))|>
  select(-contains("Population 45 to 64 years", ignore.case = TRUE))|>
  select(-contains("Population 35 to 44 years", ignore.case = TRUE))|>
  select(-contains("RACE", ignore.case = TRUE))|>
  select(-contains("college", ignore.case = TRUE))

```



```{r}
#easier code for directly deleting all the rows in a single code 
education2021 <- education2021 %>%
  select(
    -matches("Margin of Error|Total|Estimate!!Male!!|Estimate!!Female!!",
             ignore.case = TRUE)
  )
```

```{r}
# remove columns that are ONLY "(X)" (ignoring NAs)
education2021 <- education2021 %>%
  select(
    where(~ !all(.x == "(X)" | is.na(.x)))
  )
```


```{r}
#remove also percent of females and percent of males in this column (if you want), this code can be kept or deleted depending on our decision to keep Percentage Female, PErcentage Male.
education2021 <- education2021 %>%
  select(
    -matches("Estimate!!Percent Male!!|Estimate!!Percent Female!!",
             ignore.case = TRUE)
  )
```

```{r}
#convert to numeric columns the ones that need it

education2021 <- education2021 %>%
  mutate(across(
    -c(Geography, `Geographic Area Name`),
    ~ suppressWarnings(as.numeric(.x))
  ))

```

## Changes for education 2022

```{r}
#replace wrong column names with first row:
new_names_education <- as.character(education2022[1, ])
colnames(education2022) <- new_names_education
education2022 <- education2022[-1, ]
```


```{r}
#remove useless columns
education2022 <- education2022 %>%
  select(-contains("Margin of Error", ignore.case = TRUE))|>
  select(-contains("Total", ignore.case = TRUE))|>
  select(-contains("Estimate!!Male!!", ignore.case = TRUE))|>
  select(-contains("Estimate!!Female!!", ignore.case = TRUE))|>
  select(-contains("High school", ignore.case = TRUE)) |>
  select(-contains("Graduate", ignore.case = TRUE))|>
  select(-contains("18 to 24", ignore.case = TRUE))|>
  select(-contains("34", ignore.case = TRUE))|>
  select(-contains("9th", ignore.case = TRUE))|>
  select(-contains("34", ignore.case = TRUE))|>
  select(-contains("Associate's", ignore.case = TRUE))|>
  select(-contains("34", ignore.case = TRUE))|>
  select(-contains("Population 65 years and over", ignore.case = TRUE))|>
  select(-contains("Population 45 to 64 years", ignore.case = TRUE))|>
  select(-contains("Population 35 to 44 years", ignore.case = TRUE))|>
  select(-contains("RACE", ignore.case = TRUE))|>
  select(-contains("college", ignore.case = TRUE))
```



```{r}
#easier code for directly deleting all the rows in a single code 
education2022 <- education2022 %>%
  select(
    -matches("Margin of Error|Total|Estimate!!Male!!|Estimate!!Female!!",
             ignore.case = TRUE)
  )
```

```{r}
# remove columns that are ONLY "(X)" (ignoring NAs)
education2022 <- education2022 %>%
  select(
    where(~ !all(.x == "(X)" | is.na(.x)))
  )
```


```{r}
#remove also percent of females and percent of males in this column (if you want), this code can be kept or deleted depending on our decision to keep Percentage Female, PErcentage Male.
education2022 <- education2022 %>%
  select(
    -matches("Estimate!!Percent Male!!|Estimate!!Percent Female!!",
             ignore.case = TRUE)
  )
```

```{r}
#convert to numeric columns the ones that need it

education2022 <- education2022 %>%
  mutate(across(
    -c(Geography, `Geographic Area Name`),
    ~ suppressWarnings(as.numeric(.x))
  ))

```

## Changes for education 2023

```{r}
#replace wrong column names with first row:
new_names_education <- as.character(education2023[1, ])
colnames(education2023) <- new_names_education
education2023 <- education2023[-1, ]
```


```{r}
education2023 <- education2023 %>%
  select(-contains("Margin of Error", ignore.case = TRUE))|>
  select(-contains("Total", ignore.case = TRUE))|>
  select(-contains("Estimate!!Male!!", ignore.case = TRUE))|>
  select(-contains("Estimate!!Female!!", ignore.case = TRUE))|>
  select(-contains("High school", ignore.case = TRUE)) |>
  select(-contains("Graduate", ignore.case = TRUE))|>
  select(-contains("18 to 24", ignore.case = TRUE))|>
  select(-contains("34", ignore.case = TRUE))|>
  select(-contains("9th", ignore.case = TRUE))|>
  select(-contains("34", ignore.case = TRUE))|>
  select(-contains("Associate's", ignore.case = TRUE))|>
  select(-contains("34", ignore.case = TRUE))|>
  select(-contains("Population 65 years and over", ignore.case = TRUE))|>
  select(-contains("Population 45 to 64 years", ignore.case = TRUE))|>
  select(-contains("Population 35 to 44 years", ignore.case = TRUE))|>
  select(-contains("RACE", ignore.case = TRUE))|>
  select(-contains("college", ignore.case = TRUE))
```



```{r}
#easier code for directly deleting all the rows in a single code 
education2023 <- education2023 %>%
  select(
    -matches("Margin of Error|Total|Estimate!!Male!!|Estimate!!Female!!",
             ignore.case = TRUE)
  )
```

```{r}
# remove columns that are ONLY "(X)" (ignoring NAs)
education2023 <- education2023 %>%
  select(
    where(~ !all(.x == "(X)" | is.na(.x)))
  )
```


```{r}
#remove also percent of females and percent of males in this column (if you want), this code can be kept or deleted depending on our decision to keep Percentage Female, PErcentage Male.
education2023 <- education2023 %>%
  select(
    -matches("Estimate!!Percent Male!!|Estimate!!Percent Female!!",
             ignore.case = TRUE)
  )
```

```{r}
#convert to numeric columns the ones that need it

education2023 <- education2023 %>%
  mutate(across(
    -c(Geography, `Geographic Area Name`),
    ~ suppressWarnings(as.numeric(.x))
  ))

```

What is left to do for the Data cleaning part: 
- Separate the county and state in two separate columns
- Look into the issue of spec() why does it read NULL instead of specifying the columns of the data
- Remove the puerto rico variables, However this can be done just by matching the datasets
- Append the datasets to eachother so that there is only one dataset for each datasets


```{r}
Edu2121 <- 
  education2020 |>
  left_join(education2021, by = "Geographic Area Name")
```
```{r}
edu2122 <- 
  Edu2121|>
  left_join(education2022, "Geographic Area Name")
```
```{r}
edutotal <-
  edu2122|>
  left_join(education2023, "Geographic Area Name")
```
```{r}
edutotal<- edutotal|>
  select(-contains("Geography", ignore.case = TRUE))|>
  
  separate(
    col = `Geographic Area Name`,  # The column to split
    into = c("County", "State"), # The names of the two new columns
    sep = ", " # The character string used to split the column
  )
```

```{r}
# Keep only the first row encountered for each unique County name
df_unique_counties <- edutotal %>%
  distinct(County, .keep_all = TRUE)
```

```{r}
EDU2 <- df_unique_counties |>
  select(-contains("POVERTY", ignore.case = TRUE))|>
  rename(Bacmore_2020 = `Estimate!!Percent!!AGE BY EDUCATIONAL ATTAINMENT!!Population 25 years and over!!Bachelor's degree or higher.x`,
         Bacmore_2021 = `Estimate!!Percent!!AGE BY EDUCATIONAL ATTAINMENT!!Population 25 years and over!!Bachelor's degree or higher.y`,
         Bacmore_2022 = `Estimate!!Percent!!AGE BY EDUCATIONAL ATTAINMENT!!Population 25 years and over!!Bachelor's degree or higher.x.x`,
         Bacmore_2023 = `Estimate!!Percent!!AGE BY EDUCATIONAL ATTAINMENT!!Population 25 years and over!!Bachelor's degree or higher.y.y`,
         )|>
  select(-contains("Estimate", ignore.case = TRUE))
```

```{r}

EDU2Long <- EDU2 |>
  pivot_longer(
    
    cols = ends_with(c("2020", "2021", "2022", "2023")),
    names_to = c(".value", "Year"),
    
    # 3. Specify the delimiter separating the variable name and the year
    names_sep = "_",
    
    # 4. (Optional) Convert the new Year column to an integer type
    values_transform = list(Year = as.integer)
  )
```



## Description of variables

variables of Interest

- Adult smoking (in percentage of total of population of adults): This variable describes the reported rate of current daily smokers in the population of the county.

- Excessive drinking: This variable describes the self-reported rate of excessive drinkers of the population in the county.

- Teen births: This variable describes the amount of teen births per 1000 females aged 15-19.

Main variable

- Education level (percent of 25 > year olds that have a Bachelors degree minimum): This variable describes the percentage of 25 year olds or older that have a diploma for a Bachelors degree or higher. 

Control variables to complete model: 

- Unemployment: This variable describes the unemployment rate in the county. Unemployed people are 16 years old or older and are actively looking for work. 

- Income Inequality: This variable describes the income inequality levels, computed as the ratio of income at the 80th percentile over the income at the 20th percentile. 

- Severe Housing Problems: this variable describes the percentage of households that have at least one of the 4 reported housing problems: high housing costs, overcrowding, lack of kitchen or plumbing facilities.

- Children in single parent households: This variable describes the percentage of children that grow up in single parent households currently.

- Mental health providers: This variable describes the rate of mental health providers per 100000 inhabitants. 


Also need to describe the research question (500 words)


Description of the sources
The education data set comes from the United States Census Bureau. It is a national data organization that serves the mission to be the US's leading provider of data about the US citizens and the country's economy. The bureau collects a varying range of data on businesses, education, housing, 





