---
title: "Data Project"
author: "Tommaso Reali"
format: html
editor: visual
---

```{r, message= FALSE}
library(here)
library(ggplot2)
library(dplyr)
library(vroom)
library(tidyr)
```

##Dowload data

```{r}
here::i_am("Data-Management-Final-Project.Rproj")
```

Description of how to get the data sets

To get the data of the CHR&R, download each file per year seperately

Link: "https://www.countyhealthrankings.org/health-data/methodology-and-sources/data-documentation/national-data-documentation-2010-2023"

To get the education data from the census bureau: Select all of the counties as variables, pick the variable you want to pick (in this case pick the education attainment) and download the data for all the years you are interested in

Link: "https://data.census.gov/table?q=S1501:+Educational+Attainment&g=010XX00US\$0500000_040XX00US01\$0500000"

```{r, message=FALSE}
health2020 <- vroom(here("Data", "2020health.csv"))
health2021 <- vroom(here("Data", "2021health.csv"))
health2022 <- vroom(here("Data", "2022health.csv"))
health2023 <- vroom(here("Data", "2023health.csv"))
education2020 <- vroom(here("Data", "ACSST5Y2020.S1501-Data.csv"))
education2021 <- vroom(here("Data", "ACSST5Y2021.S1501-Data.csv"))
education2022 <- vroom(here("Data", "ACSST5Y2022.S1501-Data.csv"))
education2023 <- vroom(here("Data", "ACSST5Y2023.S1501-Data.csv"))

```

## Changes made to health data

## variables names to keep in dataset health:

\% Severe Housing Problems Income Ratio % Unemployed Mental Health Provider Rate Teen Birth Rate % Excessive Drinking % Smokers % Single-Parent Households

### Changes for health2020

```{r}
#replace wrong column names with first row:
new_names <- as.character(health2020[1, ])
colnames(health2020) <- new_names
health2020 <- health2020[-1, ]

```

```{r}
#remove all the columns which are not usefull (use the same technique for the other years datasets)

cols_health_2020 <- c(
  "FIPS",
  "State",
  "County",
  "% Severe Housing Problems",
  "Income Ratio",
  "% Unemployed",
  "Mental Health Provider Rate",
  "Teen Birth Rate",
  "% Excessive Drinking",
  "% Smokers",
  "% Single-Parent Households"
)

health2020 <- health2020 %>%
  select(all_of(cols_health_2020))


```

```{r}
#look at NA values and clean if they are in State, county columns
colSums(is.na(health2020))

health2020 %>%
  filter(is.na(County))

#replace NA values for county by "Average"
health2020 <- health2020 %>%
  mutate(
    County = ifelse(is.na(County), "Average", County)
  )

```

```{r}
#convert to numeric columns the ones that need it

health2020 <- health2020 %>%
  mutate(across(
    -c(FIPS, State, County),
    ~ as.numeric(gsub(",", ".", .x))
  ))

```

## Changes for Health2021

```{r}
#replace wrong column names with first row:
new_names <- as.character(health2021[1, ])
colnames(health2021) <- new_names
health2021 <- health2021[-1, ]

```

```{r}
#remove all the columns which are not usefull 

cols_health_2021 <- c(
  "FIPS",
  "State",
  "County",
  "% Severe Housing Problems",
  "Income Ratio",
  "% Unemployed",
  "Mental Health Provider Rate",
  "Teen Birth Rate",
  "% Excessive Drinking",
  "% Smokers",
  "% Children in Single-Parent Households"
)

health2021 <- health2021 %>%
  select(all_of(cols_health_2021))


```

```{r}
#look at NA values and clean if they are in State, county columns
colSums(is.na(health2021))

health2021 %>%
  filter(is.na(County))

#replace NA values for county by "Average"
health2021 <- health2021 %>%
  mutate(
    County = ifelse(is.na(County), "Average", County)
  )

```

```{r}
#convert to numeric columns the ones that need it

health2021 <- health2021 %>%
  mutate(across(
    -c(FIPS, State, County),
    ~ as.numeric(gsub(",", ".", .x))
  ))

```

## Changes for Health2022

```{r}
#replace wrong column names with first row:
new_names <- as.character(health2022[1, ])
colnames(health2022) <- new_names
health2022 <- health2022[-1, ]

```

```{r}
#remove all the columns which are not usefull 

## check if health2022 has the same variable names as 2021 and 2020

cols_health_2020 %in% names(health2022)
cols_health_2021 %in% names(health2022)

#create a vector for health2022 so we can select the columns we want

cols_health_2022 <- c(
  "FIPS",
  "State",
  "County",
  "% Severe Housing Problems",
  "Income Ratio",
  "% Unemployed",
  "Mental Health Provider Rate",
  "Teen Birth Rate",
  "% Excessive Drinking",
  "% Smokers",
  "% Children in Single-Parent Households"
)

health2022 <- health2022 %>%
  select(all_of(cols_health_2022))


```

```{r}
#look at NA values and clean if they are in State, county columns
colSums(is.na(health2022))

health2022 %>%
  filter(is.na(County))

#replace NA values for county by "Average"
health2022 <- health2022 %>%
  mutate(
    County = ifelse(is.na(County), "Average", County)
  )

```

```{r}
#convert to numeric columns the ones that need it

health2022 <- health2022 %>%
  mutate(across(
    -c(FIPS, State, County),
    ~ as.numeric(gsub(",", ".", .x))
  ))

```

## Changes for Health2023

```{r}
#replace wrong column names with first row:
new_names <- as.character(health2023[1, ])
colnames(health2023) <- new_names
health2023 <- health2023[-1, ]

```

```{r}
#remove all the columns which are not usefull 

## check if health2022 has the same variable names as 2021 and 2020

cols_health_2020 %in% names(health2023)
cols_health_2021 %in% names(health2023)
cols_health_2022 %in% names(health2023)

#create a vector for health2022 so we can select the columns we want

cols_health_2023 <- c(
  "FIPS",
  "State",
  "County",
  "% Severe Housing Problems",
  "Income Ratio",
  "% Unemployed",
  "Mental Health Provider Rate",
  "Teen Birth Rate",
  "% Excessive Drinking",
  "% Adults Reporting Currently Smoking",
  "% Children in Single-Parent Households"
)

health2023 <- health2023 %>%
  select(all_of(cols_health_2023))


```

```{r}
#look at NA values and clean if they are in State, county columns
colSums(is.na(health2023))

health2023 %>%
  filter(is.na(County))

#replace NA values for county by "Average"
health2023 <- health2023 %>%
  mutate(
    County = ifelse(is.na(County), "Average", County)
  )

```

```{r}
#convert to numeric columns the ones that need it

health2023 <- health2023 %>%
  mutate(across(
    -c(FIPS, State, County),
    ~ as.numeric(gsub(",", ".", .x))
  ))

```

## Merging the health datasets together

```{r}
#First match the names for Adult moking and single children in household

health2023 <- health2023 %>%
  rename(`% Smokers` = `% Adults Reporting Currently Smoking`)

health2020 <- health2020 %>%
  rename(`% Children in Single-Parent Households` = `% Single-Parent Households`)
```


```{r}
Hea2021 <- 
  health2020 |>
  left_join(health2021, by = c("County", "State"))
```

```{r}
Hea2122 <- 
  Hea2021|>
  left_join(health2022, by = c("County", "State"))
```

```{r}
Healthtotal <-
  Hea2122|>
  left_join(health2023, by = c("County", "State"))
```


```{r}

Healthtotal <- Healthtotal |>
  select(-contains("FIPS", ignore.case = TRUE)) |>
  rename(
    HousingP_2020   = `% Severe Housing Problems.x`,
    HousingP_2021   = `% Severe Housing Problems.y`,
    HousingP_2022   = `% Severe Housing Problems.x.x`,
    HousingP_2023   = `% Severe Housing Problems.y.y`,
    
    IncomeRatio_2020 = `Income Ratio.x`,
    IncomeRatio_2021 = `Income Ratio.y`,
    IncomeRatio_2022 = `Income Ratio.x.x`,
    IncomeRatio_2023 = `Income Ratio.y.y`,
    
    Unemployed_2020 = `% Unemployed.x`,
    Unemployed_2021 = `% Unemployed.y`,
    Unemployed_2022 = `% Unemployed.x.x`,
    Unemployed_2023 = `% Unemployed.y.y`,
    
    Mentalhealth_2020 = `Mental Health Provider Rate.x`,
    Mentalhealth_2021 = `Mental Health Provider Rate.y`,
    Mentalhealth_2022 = `Mental Health Provider Rate.x.x`,
    Mentalhealth_2023 = `Mental Health Provider Rate.y.y`,
    
    TeenBirth_2020 = `Teen Birth Rate.x`,
    TeenBirth_2021 = `Teen Birth Rate.y`,
    TeenBirth_2022 = `Teen Birth Rate.x.x`,
    TeenBirth_2023 = `Teen Birth Rate.y.y`,
    
    ExcessiveDrink_2020 = `% Excessive Drinking.x`,
    ExcessiveDrink_2021 = `% Excessive Drinking.y`,
    ExcessiveDrink_2022 = `% Excessive Drinking.x.x`,
    ExcessiveDrink_2023 = `% Excessive Drinking.y.y`,
    
    Smokers_2020 = `% Smokers.x`,
    Smokers_2021 = `% Smokers.y`,
    Smokers_2022 = `% Smokers.x.x`,
    Smokers_2023 = `% Smokers.y.y`,
    
    SingleParent_2020 = `% Children in Single-Parent Households.x`,
    SingleParent_2021 = `% Children in Single-Parent Households.y`,
    SingleParent_2022 = `% Children in Single-Parent Households.x.x`,
    SingleParent_2023 = `% Children in Single-Parent Households.y.y`
  )

```

```{r}

Healthtotal <- Healthtotal |>
  pivot_longer(
    
    cols = ends_with(c("2020", "2021", "2022", "2023")),
    names_to = c(".value", "Year"),
    
    # 3. Specify the delimiter separating the variable name and the year
    names_sep = "_",
    
    # 4. (Optional) Convert the new Year column to an integer type
    values_transform = list(Year = as.integer)
  )
```

## Changes made for Education data

###changes for education2020

```{r}
#replace wrong column names with first row:
new_names_education <- as.character(education2020[1, ])
colnames(education2020) <- new_names_education
education2020 <- education2020[-1, ]
```

```{r}
education2020 <- education2020 %>%
  select(-contains("Margin of Error", ignore.case = TRUE))|>
  select(-contains("Total", ignore.case = TRUE))|>
  select(-contains("Estimate!!Male!!", ignore.case = TRUE))|>
  select(-contains("Estimate!!Female!!", ignore.case = TRUE))|>
  select(-contains("High school", ignore.case = TRUE)) |>
  select(-contains("Graduate", ignore.case = TRUE))|>
  select(-contains("18 to 24", ignore.case = TRUE))|>
  select(-contains("34", ignore.case = TRUE))|>
  select(-contains("9th", ignore.case = TRUE))|>
  select(-contains("34", ignore.case = TRUE))|>
  select(-contains("Associate's", ignore.case = TRUE))|>
  select(-contains("34", ignore.case = TRUE))|>
  select(-contains("Population 65 years and over", ignore.case = TRUE))|>
  select(-contains("Population 45 to 64 years", ignore.case = TRUE))|>
  select(-contains("Population 35 to 44 years", ignore.case = TRUE))|>
  select(-contains("RACE", ignore.case = TRUE))|>
  select(-contains("college", ignore.case = TRUE))
```

```{r}
#easier code for directly deleting all the rows in a single code 
education2020 <- education2020 %>%
  select(
    -matches("Margin of Error|Total|Estimate!!Male!!|Estimate!!Female!!",
             ignore.case = TRUE)
  )
```

```{r}
# remove columns that are ONLY "(X)" (ignoring NAs)
education2020 <- education2020 %>%
  select(
    where(~ !all(.x == "(X)" | is.na(.x)))
  )
```

```{r}
#remove also percent of females and percent of males in this column (if you want), this code can be kept or deleted depending on our decision to keep Percentage Female, PErcentage Male.
education2020 <- education2020 %>%
  select(
    -matches("Estimate!!Percent Male!!|Estimate!!Percent Female!!",
             ignore.case = TRUE)
  )
```

```{r}
#convert to numeric columns the ones that need it

education2020 <- education2020 %>%
  mutate(across(
    -c(Geography, `Geographic Area Name`),
    ~ suppressWarnings(as.numeric(.x))
  ))

```

## Changes for education 2021

```{r}
#replace wrong column names with first row:
new_names_education <- as.character(education2021[1, ])
colnames(education2021) <- new_names_education
education2021 <- education2021[-1, ]
```

```{r}
education2021 <- education2021 %>%
  select(-contains("Margin of Error", ignore.case = TRUE))|>
  select(-contains("Total", ignore.case = TRUE))|>
  select(-contains("Estimate!!Male!!", ignore.case = TRUE))|>
  select(-contains("Estimate!!Female!!", ignore.case = TRUE))|>
  select(-contains("High school", ignore.case = TRUE)) |>
  select(-contains("Graduate", ignore.case = TRUE))|>
  select(-contains("18 to 24", ignore.case = TRUE))|>
  select(-contains("34", ignore.case = TRUE))|>
  select(-contains("9th", ignore.case = TRUE))|>
  select(-contains("34", ignore.case = TRUE))|>
  select(-contains("Associate's", ignore.case = TRUE))|>
  select(-contains("34", ignore.case = TRUE))|>
  select(-contains("Population 65 years and over", ignore.case = TRUE))|>
  select(-contains("Population 45 to 64 years", ignore.case = TRUE))|>
  select(-contains("Population 35 to 44 years", ignore.case = TRUE))|>
  select(-contains("RACE", ignore.case = TRUE))|>
  select(-contains("college", ignore.case = TRUE))

```

```{r}
#easier code for directly deleting all the rows in a single code 
education2021 <- education2021 %>%
  select(
    -matches("Margin of Error|Total|Estimate!!Male!!|Estimate!!Female!!",
             ignore.case = TRUE)
  )
```

```{r}
# remove columns that are ONLY "(X)" (ignoring NAs)
education2021 <- education2021 %>%
  select(
    where(~ !all(.x == "(X)" | is.na(.x)))
  )
```

```{r}
#remove also percent of females and percent of males in this column (if you want), this code can be kept or deleted depending on our decision to keep Percentage Female, PErcentage Male.
education2021 <- education2021 %>%
  select(
    -matches("Estimate!!Percent Male!!|Estimate!!Percent Female!!",
             ignore.case = TRUE)
  )
```

```{r}
#convert to numeric columns the ones that need it

education2021 <- education2021 %>%
  mutate(across(
    -c(Geography, `Geographic Area Name`),
    ~ suppressWarnings(as.numeric(.x))
  ))

```

## Changes for education 2022

```{r}
#replace wrong column names with first row:
new_names_education <- as.character(education2022[1, ])
colnames(education2022) <- new_names_education
education2022 <- education2022[-1, ]
```

```{r}
#remove useless columns
education2022 <- education2022 %>%
  select(-contains("Margin of Error", ignore.case = TRUE))|>
  select(-contains("Total", ignore.case = TRUE))|>
  select(-contains("Estimate!!Male!!", ignore.case = TRUE))|>
  select(-contains("Estimate!!Female!!", ignore.case = TRUE))|>
  select(-contains("High school", ignore.case = TRUE)) |>
  select(-contains("Graduate", ignore.case = TRUE))|>
  select(-contains("18 to 24", ignore.case = TRUE))|>
  select(-contains("34", ignore.case = TRUE))|>
  select(-contains("9th", ignore.case = TRUE))|>
  select(-contains("34", ignore.case = TRUE))|>
  select(-contains("Associate's", ignore.case = TRUE))|>
  select(-contains("34", ignore.case = TRUE))|>
  select(-contains("Population 65 years and over", ignore.case = TRUE))|>
  select(-contains("Population 45 to 64 years", ignore.case = TRUE))|>
  select(-contains("Population 35 to 44 years", ignore.case = TRUE))|>
  select(-contains("RACE", ignore.case = TRUE))|>
  select(-contains("college", ignore.case = TRUE))
```

```{r}
#easier code for directly deleting all the rows in a single code 
education2022 <- education2022 %>%
  select(
    -matches("Margin of Error|Total|Estimate!!Male!!|Estimate!!Female!!",
             ignore.case = TRUE)
  )
```

```{r}
# remove columns that are ONLY "(X)" (ignoring NAs)
education2022 <- education2022 %>%
  select(
    where(~ !all(.x == "(X)" | is.na(.x)))
  )
```

```{r}
#remove also percent of females and percent of males in this column (if you want), this code can be kept or deleted depending on our decision to keep Percentage Female, PErcentage Male.
education2022 <- education2022 %>%
  select(
    -matches("Estimate!!Percent Male!!|Estimate!!Percent Female!!",
             ignore.case = TRUE)
  )
```

```{r}
#convert to numeric columns the ones that need it

education2022 <- education2022 %>%
  mutate(across(
    -c(Geography, `Geographic Area Name`),
    ~ suppressWarnings(as.numeric(.x))
  ))

```

## Changes for education 2023

```{r}
#replace wrong column names with first row:
new_names_education <- as.character(education2023[1, ])
colnames(education2023) <- new_names_education
education2023 <- education2023[-1, ]
```

```{r}
education2023 <- education2023 %>%
  select(-contains("Margin of Error", ignore.case = TRUE))|>
  select(-contains("Total", ignore.case = TRUE))|>
  select(-contains("Estimate!!Male!!", ignore.case = TRUE))|>
  select(-contains("Estimate!!Female!!", ignore.case = TRUE))|>
  select(-contains("High school", ignore.case = TRUE)) |>
  select(-contains("Graduate", ignore.case = TRUE))|>
  select(-contains("18 to 24", ignore.case = TRUE))|>
  select(-contains("34", ignore.case = TRUE))|>
  select(-contains("9th", ignore.case = TRUE))|>
  select(-contains("34", ignore.case = TRUE))|>
  select(-contains("Associate's", ignore.case = TRUE))|>
  select(-contains("34", ignore.case = TRUE))|>
  select(-contains("Population 65 years and over", ignore.case = TRUE))|>
  select(-contains("Population 45 to 64 years", ignore.case = TRUE))|>
  select(-contains("Population 35 to 44 years", ignore.case = TRUE))|>
  select(-contains("RACE", ignore.case = TRUE))|>
  select(-contains("college", ignore.case = TRUE))
```

```{r}
#easier code for directly deleting all the rows in a single code 
education2023 <- education2023 %>%
  select(
    -matches("Margin of Error|Total|Estimate!!Male!!|Estimate!!Female!!",
             ignore.case = TRUE)
  )
```

```{r}
# remove columns that are ONLY "(X)" (ignoring NAs)
education2023 <- education2023 %>%
  select(
    where(~ !all(.x == "(X)" | is.na(.x)))
  )
```

```{r}
#remove also percent of females and percent of males in this column (if you want), this code can be kept or deleted depending on our decision to keep Percentage Female, PErcentage Male.
education2023 <- education2023 %>%
  select(
    -matches("Estimate!!Percent Male!!|Estimate!!Percent Female!!",
             ignore.case = TRUE)
  )
```

```{r}
#convert to numeric columns the ones that need it

education2023 <- education2023 %>%
  mutate(across(
    -c(Geography, `Geographic Area Name`),
    ~ suppressWarnings(as.numeric(.x))
  ))

```

What is left to do for the Data cleaning part: - Separate the county and state in two separate columns - Look into the issue of spec() why does it read NULL instead of specifying the columns of the data - Remove the puerto rico variables, However this can be done just by matching the datasets - Append the datasets to eachother so that there is only one dataset for each datasets

```{r}
Edu2121 <- 
  education2020 |>
  left_join(education2021, by = "Geographic Area Name")
```

```{r}
edu2122 <- 
  Edu2121|>
  left_join(education2022, "Geographic Area Name")
```

```{r}
edutotal <-
  edu2122|>
  left_join(education2023, "Geographic Area Name")
```

```{r}
edutotal<- edutotal|>
  select(-contains("Geography", ignore.case = TRUE))|>
  
  separate(
    col = `Geographic Area Name`,  # The column to split
    into = c("County", "State"), # The names of the two new columns
    sep = ", " # The character string used to split the column
  )
```

```{r}
# Keep only the first row encountered for each unique County name
df_unique_counties <- edutotal %>%
  distinct(County, .keep_all = TRUE)
```

```{r}
EDU2 <- df_unique_counties |>
  select(-contains("POVERTY", ignore.case = TRUE))|>
  rename(Bacmore_2020 = `Estimate!!Percent!!AGE BY EDUCATIONAL ATTAINMENT!!Population 25 years and over!!Bachelor's degree or higher.x`,
         Bacmore_2021 = `Estimate!!Percent!!AGE BY EDUCATIONAL ATTAINMENT!!Population 25 years and over!!Bachelor's degree or higher.y`,
         Bacmore_2022 = `Estimate!!Percent!!AGE BY EDUCATIONAL ATTAINMENT!!Population 25 years and over!!Bachelor's degree or higher.x.x`,
         Bacmore_2023 = `Estimate!!Percent!!AGE BY EDUCATIONAL ATTAINMENT!!Population 25 years and over!!Bachelor's degree or higher.y.y`,
         )|>
  select(-contains("Estimate", ignore.case = TRUE))
```

```{r}

EDU2Long <- EDU2 |>
  pivot_longer(
    
    cols = ends_with(c("2020", "2021", "2022", "2023")),
    names_to = c(".value", "Year"),
    
    # 3. Specify the delimiter separating the variable name and the year
    names_sep = "_",
    
    # 4. (Optional) Convert the new Year column to an integer type
    values_transform = list(Year = as.integer)
  )
```

## Description of variables

variables of Interest

-   Adult smoking (in percentage of total of population of adults): This variable describes the reported rate of current daily smokers in the population of the county.

-   Excessive drinking: This variable describes the self-reported rate of excessive drinkers of the population in the county.

-   Teen births: This variable describes the amount of teen births per 1000 females aged 15-19.

Main variable

-   Education level (percent of 25 \> year olds that have a Bachelors degree minimum): This variable describes the percentage of 25 year olds or older that have a diploma for a Bachelors degree or higher.

Control variables to complete model:

-   Unemployment: This variable describes the unemployment rate in the county. Unemployed people are 16 years old or older and are actively looking for work.

-   Income Inequality: This variable describes the income inequality levels, computed as the ratio of income at the 80th percentile over the income at the 20th percentile.

-   Severe Housing Problems: this variable describes the percentage of households that have at least one of the 4 reported housing problems: high housing costs, overcrowding, lack of kitchen or plumbing facilities.

-   Children in single parent households: This variable describes the percentage of children that grow up in single parent households currently.

-   Mental health providers: This variable describes the rate of mental health providers per 100000 inhabitants.

Also need to describe the research question (500 words)


## Description of the data joining and cleaning

The raw data for health was at first downloaded from the CHR&R website for all the years we are interested in for this study (2020-2023) as 4 different csv files (one for each year). After downloading the data in Rstudio using vroom(), the the raw columns names were replaced with the true column names which were situated in the first row. Furthermore, all the columns which were not useful were removed in order to keep only the variables used in this study, namely: % Severe Housing Problems, Income Ratio, % Unemployed, Mental Health Provider Rate, Teen Birth Rate, % Excessive Drinking, % Smokers, % Children in Single-Parent Households. Moreover, the missing values for county names were replaced by "Average" as for each county there was one observation per year calculating its average in the raw dataset. The nature of the variables was also changed: county and state are characters, and the rest are numerical. Finally, after having cleaned the 4 years datasets, they were merged together with the column (state, county), ensuring every state and county are the correct key. After the merge the dataset was pivoted to be in long format (with each observation a unique combination of State, county and year), and the columns were renamed as needed. 

## Description of the sources

The Health dataset used in this study is constructed of data coming entirely from the CHR&R (County Health Rankings and Roadmaps) program. This initiative was created by the University of Wisconsin Population Health institute with support from the Robert Wood Johnson Foundation. It consists in providing annual data on health factors within and across almost the entirety of counties from the United States. The CHR&R provides data, evidence, guidance and examples to build awareness about the multiple factors influencing health conditions and disparities across the country. Furthermore, CHR&R puts an emphasis on the range of health-influencing measures that offer crucial context about community specific health conditions. In fact, Along with the data the program provides tools and evidence-based strategies, which can be used by the communities to move to action and solve the important health and societal issues concerning their people. As an example, The "2025 Annual data release" includes over 80 health-related measures, and focuses on the impact of certain societal factors present in the United States on community conditions.

## Table of health data set

```{r}
data_name <- "Health Data"

row_count <- nrow(Healthtotal)
column_count <- ncol(Healthtotal)

summary_table <- data.frame(
  `Data File Name` = data_name,
  `Number of Rows` = row_count,
  `Number of Columns` = column_count,
  check.names = FALSE
)
print(summary_table)
```

## Table for health main target variables (% Smokers, % Excessive drinkers, Teen birth rates)

Since there are 3144 counties, the data will be shown as the aggregated average of the country per year.

```{r, message=FALSE}
Healthtotal|>
  ggplot(aes(y = TeenBirth, x = Year)) +
  geom_violin()
```

These violin plots show the aggregated US average of Teen birth rates as the amount of teen births per 1000 females aged 15-19. It is observed that the density of the violin plots gets more and more concentrated towards the bottom as the years pass by. As a matter of fact, the average of aggregated teen birth lowers from about 25% in 2020 to 19% in 2023.

```{r, message=FALSE}
Healthtotal|>
  ggplot(aes(y = Smokers, x = Year)) +
  geom_violin()
```

These violin plots show the aggregated US average of adult smokers as a percentage of the population. It can be seen that each year the average of percentage of smokers gradually increases and the density of the violin plots gets slimmer. In fact, in 2020 the average percentage of aggregated smokers in the US is about just above 15%. It grows up until nearly 20% in 2021, and decreases to about 18% in 2023. 

```{r, message=FALSE}
Healthtotal|>
  ggplot(aes(y = ExcessiveDrink, x = Year)) +
  geom_violin()
```


These violin plots showing the aggregated US average of Excessive drinking as a percentage of the population highlight a thicker upper tail as the years pass by, meaning there is a higher count of people excessively drinking. However the average of the percentage of excessive drinkers doesn't seem to change from from 2020 to 2023.